R code for 'Apparent physiological costs of blood parasite infection in the early-life of a vertebrate host - only in highest infections

### 2022-02-24 MS_physiology_infection.Rmd
Add files via upload

##  Create Figure 2a
## -----------------------------------------------------------------------------
Fig_2A  <- ggplot(tab_cont) +  
  aes(Timepoint, Count, fill = Prevalence) +  
  facet_wrap(facets = "Treatment") +  
  geom_bar(stat = 'identity',  width = 0.3) +  
  labs(y = "Count of nestlings") +  
  geom_text(aes(label = Count),position  = position_stack(0.5)) +  
  scale_fill_manual(values = c("#D0D0D0","#A8A8A8")) +  
  theme(axis.title.x  =  element_blank(), 
        strip.text = element_text(size = 16, face = "bold"),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 16),
        axis.title.y = element_text(size = 18, margin =  unit(c(0, 3, 0, 0), "mm")),
        axis.text = element_text(size = 16))
Fig_2A

@@

perm.t.test(as.numeric(as.character(df$Infection_intensity.1))~df$Treatment)

@@

No treatment effect on breathing rate!

@@

# Plotting predicted values across infection intensities
## -----------------------------------------------------------------------------
Fig_5A = ggplot(mod_br_para) + 
  aes(as.factor(mod_br_para$x),mod_br_para$predicted) + 
  geom_hline(yintercept = mean(mod_br_para$predicted, na.rm = T), linetype = "dashed") + 
  labs(x = "", y = "Breathing rate") + 
  geom_errorbar(aes(ymin = mod_br_para$lower_CI,
                    ymax = mod_br_para$higher_CI),colour = "black",size = 2, width = .1) + 
  geom_errorbar(aes(ymin = mod_br_para$predicted - mod_br_para$se, 
                    ymax = mod_br_para$predicted + mod_br_para$se),
                size = 8, colour = wes_palette("Zissou1",5), width = 0,) + 
  geom_point(color = "black", size = 4) + 
  theme_classic() + 
  theme(legend.position = "none",
        axis.title = element_text(size = 14),
        axis.text = element_text(size = 12),
        axis.line.x = element_blank(),
        axis.ticks.x = element_blank(), plot.margin = margin(0, 0, 0, 0, "cm"),
        plot.title = element_text(hjust = 0.5, face = "bold")) + 
  geom_segment(size = .8, aes(x = 1, y = max(mod_br_para$higher_CI, na.rm = T) + 1.2,
                              xend = 5, yend = max(mod_br_para$higher_CI, na.rm = T) + 1.2)) + 
    geom_text(aes(x = 3, y = max(mod_br_para$higher_CI, na.rm = T) + 2),
              label = "n.s.", size = 5)
```

@@

## Plotting predicted values across infection intensities
## -----------------------------------------------------------------------------
Fig_5B = ggplot(mod_bt_para) + 
  aes(as.factor(mod_bt_para$x),mod_bt_para$predicted) + 
  geom_hline(yintercept = mean(mod_bt_para$predicted, na.rm = T), linetype = "dashed") + 
  geom_errorbar(aes(ymin = mod_bt_para$lower_CI,ymax = mod_bt_para$higher_CI),colour = "black",size = 2, width = .1) + 
  geom_errorbar(aes(ymin = mod_bt_para$predicted - mod_bt_para$se,ymax = mod_bt_para$predicted + mod_bt_para$se),
                size = 8,colour = wes_palette("Zissou1",5), width = 0,) + 
  labs(x = "", y = "Body temperature") + 
  geom_point(color = "black", size = 4) + 
  theme_classic() + 
  theme(legend.position = "none",
        axis.title = element_text(size = 14),
        axis.text = element_text(size = 12),
        axis.line.x = element_blank(), 
        axis.ticks.x = element_blank(), plot.margin = margin(0, 0, 0, 0, "cm"),
        plot.title = element_text(hjust = 0.5, face = "bold")) + 
  geom_segment(size = .8, aes(x = 1, y = max(mod_bt_para$higher_CI, na.rm = T) + 0.2,
                              xend = 5, yend = max(mod_bt_para$higher_CI, na.rm = T) + 0.2)) + 
  geom_text(aes(x = 3, y = max(mod_bt_para$higher_CI, na.rm = T) + 0.25),
            label = "0.006", size = 5) + 
  geom_segment(size = .8,
               aes(x = 1, y = max(mod_bt_para$higher_CI, na.rm = T) + 0.2,
                   xend = 1, yend = max(mod_bt_para$higher_CI, na.rm = T) + 0.05)) + 
  geom_segment(size = .8,
               aes(x = 5, y = max(mod_bt_para$higher_CI, na.rm = T) + 0.05,
                   xend = 5, yend = max(mod_bt_para$higher_CI, na.rm = T) + 0.2))
```

@@

Fig_5C = ggplot(mod_bc_para) + 
  aes(as.factor(mod_bc_para$x),mod_bc_para$predicted) + 
  geom_hline(yintercept = mean(mod_bc_para$predicted, na.rm = T), linetype = "dashed") + 
  labs(x = "", y = "Δ Body condition") + 
  geom_errorbar(aes(ymin = mod_bc_para$lower_CI, ymax = mod_bc_para$higher_CI),
                colour = "black",size = 2, width = .1) + 
  geom_errorbar(aes(ymin = mod_bc_para$predicted - mod_bc_para$se,
                    ymax = mod_bc_para$predicted + mod_bc_para$se),
                size = 8, colour = wes_palette("Zissou1",5), width = 0,) + 
  geom_point(color = "black", size = 4) + 
  theme_classic() + 
  theme(legend.position = "none",
        axis.title = element_text(size = 14),  
        axis.text = element_text(size = 12), 
        axis.line.x = element_blank(),
        axis.ticks.x = element_blank(), plot.margin = margin(0, 0, 0, 0, "cm"),
        plot.title = element_text(hjust = 0.5, face = "bold")) + 
  geom_segment(size = .8,
                 aes(x = 1, y = -105, xend = 5, yend = -105)) + 
    geom_text(aes(x = 3, y = -110),
              label = "p=0.08", size = 5) + 
  geom_segment(size = .8,
                 aes(x = 1, y = -105, xend = 1, yend = -105 + 5)) + 
  geom_segment(size = .8,
                 aes(x = 5, y = -105, xend = 5, yend = -105 + 5))
```

@@

# Plotting predicted values across infection intensities
## -----------------------------------------------------------------------------
Fig_5D = ggplot(mod_gr_para) + 
  aes(as.factor(mod_gr_para$x),mod_gr_para$predicted) + 
  geom_hline(yintercept = mean(mod_gr_para$predicted, na.rm = T), linetype = "dashed") + 
  labs(x = "", y = "Growth rate") + 
  geom_errorbar(aes(ymin = mod_gr_para$lower_CI,
                    ymax = mod_gr_para$higher_CI), colour = "black",size = 2, width = .1) + 
  geom_errorbar(aes(ymin = mod_gr_para$predicted - mod_gr_para$se,
                    ymax = mod_gr_para$predicted + mod_gr_para$se),size = 8, 
                colour = wes_palette("Zissou1",5), width = 0) + 
  geom_point(color = "black", outlier.shape = NA, size = 4, width = .3) + 
  theme_classic() + 
  theme(legend.position = "none", 
        axis.title = element_text(size = 14),
        axis.text = element_text(size = 12),
        axis.line.x = element_blank(),
        axis.ticks.x = element_blank(), plot.margin = margin(0, 0, 0, 0, "cm"),
        plot.title = element_text(hjust = 0.5, face = "bold")) + 
  geom_segment(size = .8,
               aes(x = 1, y = 4, xend = 5, yend = 4)) + 
  geom_text(aes(x = 3, y = 5 - 0.63),
            label = "p = 0.094", size = 5) + 
  geom_segment(size = .8,
               aes(x = 1, y = 4, xend = 1, yend = 4 + 0.5)) + 
  geom_segment(size = .8,
               aes(x = 5, y = 4, xend = 5, yend = 4 + 0.5))
```

@@

## Modeling host health to Infection intensity
## -----------------------------------------------------------------------------
tolres_mod = lm(delta_body_condition~Infection_intensity.2, data = tol_resdata)
summary(tolres_mod)

@@

## Comparison of slopes to zero
## -----------------------------------------------------------------------------
tolre_deltasum = summary(lms_delta)
summary(lms_delta, infer = T)

@@

## Arrange plot for three infection stages 
## -----------------------------------------------------------------------------
g <- ggplot_gtable(ggplot_build(tolres))
stripr <- which(grepl('strip-t', g$layout$name))
fills <- c("#3B9AB2", "#E1AF00", "#F21A00")
k <- 1
for (i in stripr) {
 j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
k <- k + 1
}

>>>>

##  Create Figure 2a
## -----------------------------------------------------------------------------
Fig_2A  <- ggplot(tab_cont) +  
  aes(Timepoint, Count, fill = Prevalence) +  
  facet_wrap(facets = "Treatment") +  
  geom_bar(stat = 'identity',  width = 0.3) +  
  labs(y = "Count of nestlings") +  
  geom_text(aes(label = Count),position  = position_stack(0.5)) +  
  scale_fill_manual(values = c("#D0D0D0","#A8A8A8")) +  
  theme(axis.title.x  =  element_blank(), 
        strip.text = element_text(size = 16, face = "bold"),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 16),
        axis.title.y = element_text(size = 18, margin =  unit(c(0, 3, 0, 0), "mm")),
        axis.text = element_text(size = 16))

@@

perm.t.test(as.numeric(as.character(df$Infection_intensity.1))~df$Treatment, progress = F)

@@

No treatment effect on breathing rate

@@

# Plotting predicted values across infection intensities
## -----------------------------------------------------------------------------
Fig_5A = ggplot(mod_br_para) + 
  aes(as.factor(mod_br_para$x),mod_br_para$predicted) + 
  geom_hline(yintercept = mean(mod_br_para$predicted, na.rm = T), linetype = "dashed") + 
  labs(x = "", y = "Breathing rate") + 
  geom_errorbar(aes(ymin = mod_br_para$lower_CI,
                    ymax = mod_br_para$higher_CI),colour = "black",size = 2, width = .1) + 
  geom_errorbar(aes(ymin = mod_br_para$predicted - mod_br_para$se, 
                    ymax = mod_br_para$predicted + mod_br_para$se),
                size = 8, colour = wes_palette("Zissou1",5), width = 0,) + 
  geom_point(color = "black", size = 4) + 
  theme_classic() + 
  theme(legend.position = "none",
        axis.title = element_text(size = 14),
        axis.text = element_text(size = 12),
        axis.line.x = element_blank(),
        axis.ticks.x = element_blank(), plot.margin = margin(0, 0, 0, 0, "cm"),
        plot.title = element_text(hjust = 0.5, face = "bold")) + 
  geom_segment(size = .8, aes(x = 1, y = max(mod_br_para$higher_CI, na.rm = T) + 1.2,
                              xend = 5, yend = max(mod_br_para$higher_CI, na.rm = T) + 1.2)) + 
    geom_text(aes(x = 3, y = max(mod_br_para$higher_CI, na.rm = T) + 2),
              label = "n.s.", size = 6)
```

@@

## Plotting predicted values across infection intensities
## -----------------------------------------------------------------------------
Fig_5B = ggplot(mod_bt_para) + 
  aes(as.factor(mod_bt_para$x),mod_bt_para$predicted) + 
  geom_hline(yintercept = mean(mod_bt_para$predicted, na.rm = T), linetype = "dashed") + 
  geom_errorbar(aes(ymin = mod_bt_para$lower_CI,ymax = mod_bt_para$higher_CI),colour = "black",size = 2, width = .1) + 
  geom_errorbar(aes(ymin = mod_bt_para$predicted - mod_bt_para$se,ymax = mod_bt_para$predicted + mod_bt_para$se),
                size = 8,colour = wes_palette("Zissou1",5), width = 0,) + 
  labs(x = "", y = "Body temperature") + 
  geom_point(color = "black", size = 4) + 
  theme_classic() + 
  theme(legend.position = "none",
        axis.title = element_text(size = 14),
        axis.text = element_text(size = 12),
        axis.line.x = element_blank(), 
        axis.ticks.x = element_blank(), plot.margin = margin(0, 0, 0, 0, "cm"),
        plot.title = element_text(hjust = 0.5, face = "bold")) + 
  geom_segment(size = .8, aes(x = 1, y = max(mod_bt_para$higher_CI, na.rm = T) + 0.2,
                              xend = 5, yend = max(mod_bt_para$higher_CI, na.rm = T) + 0.2)) + 
  geom_text(aes(x = 3, y = max(mod_bt_para$higher_CI, na.rm = T) + 0.3),
            label = "P = 0.006", size = 5) + 
  geom_segment(size = .8,
               aes(x = 1, y = max(mod_bt_para$higher_CI, na.rm = T) + 0.2,
                   xend = 1, yend = max(mod_bt_para$higher_CI, na.rm = T) + 0.1)) + 
  geom_segment(size = .8,
               aes(x = 5, y = max(mod_bt_para$higher_CI, na.rm = T) + 0.1,
                   xend = 5, yend = max(mod_bt_para$higher_CI, na.rm = T) + 0.2))
```

@@

Fig_5C = ggplot(mod_bc_para) + 
  aes(as.factor(mod_bc_para$x),mod_bc_para$predicted) + 
  geom_hline(yintercept = mean(mod_bc_para$predicted, na.rm = T), linetype = "dashed") + 
  labs(x = "", y = "Δ Body condition") + 
  geom_errorbar(aes(ymin = mod_bc_para$lower_CI, ymax = mod_bc_para$higher_CI),
                colour = "black",size = 2, width = .1) + 
  geom_errorbar(aes(ymin = mod_bc_para$predicted - mod_bc_para$se,
                    ymax = mod_bc_para$predicted + mod_bc_para$se),
                size = 8, colour = wes_palette("Zissou1",5), width = 0,) + 
  geom_point(color = "black", size = 4) + 
  theme_classic() + 
  theme(legend.position = "none",
        axis.title = element_text(size = 14),  
        axis.text = element_text(size = 12), 
        axis.line.x = element_blank(),
        axis.ticks.x = element_blank(), plot.margin = margin(0, 0, 0, 0, "cm"),
        plot.title = element_text(hjust = 0.5, face = "bold")) + 
  geom_segment(size = .8,
                 aes(x = 1, y = 60, xend = 5, yend = 60)) + 
    geom_text(aes(x = 3, y = 67),
              label = "P = 0.080", size = 5) + 
  geom_segment(size = .8,
                 aes(x = 1, y = 60, xend = 1, yend = 53)) + 
  geom_segment(size = .8,
                 aes(x = 5, y = 60, xend = 5, yend = 53))
```

@@

# Plotting predicted values across infection intensities
## -----------------------------------------------------------------------------
Fig_5D = ggplot(mod_gr_para) + 
  aes(as.factor(mod_gr_para$x),mod_gr_para$predicted) + 
  geom_hline(yintercept = mean(mod_gr_para$predicted, na.rm = T), linetype = "dashed") + 
  labs(x = "", y = "Growth rate") + 
  geom_errorbar(aes(ymin = mod_gr_para$lower_CI,
                    ymax = mod_gr_para$higher_CI), colour = "black",size = 2, width = .1) + 
  geom_errorbar(aes(ymin = mod_gr_para$predicted - mod_gr_para$se,
                    ymax = mod_gr_para$predicted + mod_gr_para$se),size = 8, 
                colour = wes_palette("Zissou1",5), width = 0) + 
  geom_point(color = "black", outlier.shape = NA, size = 4, width = .3) + 
  theme_classic() + 
  theme(legend.position = "none", 
        axis.title = element_text(size = 14),
        axis.text = element_text(size = 12),
        axis.line.x = element_blank(),
        axis.ticks.x = element_blank(), plot.margin = margin(0, 0, 0, 0, "cm"),
        plot.title = element_text(hjust = 0.5, face = "bold")) + 
  geom_segment(size = .8,
               aes(x = 1, y = 30, xend = 5, yend = 30)) + 
  geom_text(aes(x = 3, y = 31),
            label = "P = 0.094", size = 5) + 
  geom_segment(size = .8,
               aes(x = 1, y = 29, xend = 1, yend = 30)) + 
  geom_segment(size = .8,
               aes(x = 5, y = 29, xend = 5, yend = 30))
```

@@

## Modeling host health to Infection intensity
## -----------------------------------------------------------------------------
lm(delta_body_condition~Infection_intensity.2, data = tol_resdata) %>%
summary()

@@

## Comparison of slopes to zero
## -----------------------------------------------------------------------------

summary(lms_delta, infer = T)

@@

## Arrange plot for three infection stages 
## -----------------------------------------------------------------------------
g <- ggplot_gtable(ggplot_build(tolres))
stripr <- which(grepl('strip-t', g$layout$name))
fills <- c("#4CC6E5", "#F7C600", "#FF5643")
k <- 1
for (i in stripr) {
 j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
k <- k + 1
}
